---
title: "Problem Specification"
# author: "katerobsau"
# date: "2019-08-05"
output: workflowr::wflow_html
editor_options:
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(cowplot)
library(knitr)
knitr::opts_knit$set(eval.after = 'fig.cap')
source("code/figure_labels.R")
```

<style>
p.comment {
background-color: #DBDBDB;
padding: 10px;
border: 1px solid black;
margin-left: 25px;
border-radius: 5px;
font-style: italic;
}
</style>

## Compound Events

Individually, both extreme rainfall and storm surge have the potential to produce flooding. However,compared with either event occurring individual, the potential risk is magnified if these events co-occur. This is what is known as a compound event.

<p class="comment">
**Definition 1:** The IPCC SREX (Ref 7, p. 118) defines a compound event as:  
(1) two or more extreme events occurring simultaneously or successively,   
(2) combinations of extreme events with underlying conditions that amplify the impact of the events, or   
(3) combinations of events that are not themselves extremes but lead  
to an extreme event or impact when combined.  
The contributing events can be of similar (clustered multiple events) or different type(s).
</p>

<!-- <div class="alert alert-info"> -->
<!--   Different box example -->
<!-- </div> -->

## Rainfall-Surge Mechansims

Wahl et al. (2015) breaks the compound event involving rainfall-surge down into three mechanism types. 

<p class="comment">
**Mechanism (1)** in estuarine regions, the joint occurrence of both may elevate water levels to a point where flooding is initiated or its impacts exacerbated.   
**Mechanism (2)** occurs when a destructive storm surge already causes widespread flooding, such that any significant rainfall on top of this—even if it is not an extreme event on its own—increases the flood depth and/or extent of the inundated area.   
**Mechanism (3)** occurs during a moderate storm surge which does not directly cause flooding, but is high enough to fully block or slow down gravity-fed storm water drainage, such that precipitation is more likely to cause flooding.
</p>

An example of the risk regions related to these three mechanisms is shown in Figure `r figRef("riskregions")`.

```{r region a, echo = FALSE}
delta = 0.75
regions <- data.frame(
  Risk = rep(c("M3: Rain-only", "M1: Both", "M2: Surge-only"), each = 4),
  x = c(c(0, delta, delta, 0), 
        c(delta, 1, 1, delta), 
        c(delta, 1, 1, delta)),
  y = c(c(delta, delta, 1, 1),
        c(delta, delta, 1, 1),
        c(0, 0, delta, delta))
)

region_a <- ggplot() + 
  geom_polygon(data = regions, aes(x = x, y = y, fill = Risk), alpha = 0.75) +
  geom_hline(yintercept = delta) + 
  geom_vline(xintercept = delta) + 
  xlim(c(0,1)) +
  xlab("Surge Height") + 
  ylab("Rainfall Amount") +
  ylim(c(0,1)) + 
  theme_bw() +
  theme(#axis.title = element_blank(),
        axis.text = element_blank(), 
        legend.position = c(0.3, 0.2))
# region_a
```

```{r region b, echo = FALSE}
delta = 0.8
theta = seq(0, pi/2, length.out = 90)
theta_rain = theta[theta >= pi/2*3/5]
theta_surge = theta[theta <= pi/2*2/5]
theta_both = setdiff(theta, c(theta_rain, theta_surge))

x_rain = c(delta*cos(theta_rain), 0, cos(min(theta_rain)))
y_rain = c(delta*sin(theta_rain), 1, 1)
x_surge = c(delta*cos(theta_surge), 1, 1)
y_surge = c(delta*sin(theta_surge), sin(max(theta_surge)), 0)
x_both = c(delta*cos(theta_both), cos(max(theta_both)), 1, 1)
y_both = c(delta*sin(theta_both), 1, 1, sin(min(theta_both)))
# plot(x_both, y_both)
            
regions_circle <- data.frame(
  Risk = rep(c("M3: Rain-only", "M1: Both", "M2: Surge-only"), 
             times = c(length(x_rain),length(x_both),length(x_surge))),
  x = c(x_rain, x_both, x_surge),
  y = c(y_rain, y_both, y_surge)
)

region_b <- ggplot() + 
  geom_polygon(data = regions_circle, aes(x = x, y = y, fill = Risk), alpha = 0.75) +
  geom_line(data = NULL, aes(x = delta*cos(theta), y = delta*sin(theta))) + 
  geom_line(data = NULL, aes(x = c(delta*cos(min(theta_rain)), cos(min(theta_rain))),
                             y = c(delta*sin(min(theta_rain)) , 1))) +
  geom_line(data = NULL, aes(x = c(delta*cos(max(theta_surge)), 1),
                             y = c(delta*sin(max(theta_surge)) , sin(max(theta_surge))))) +
  xlim(c(0,1)) +
  ylim(c(0,1)) + 
  xlab("Surge Height") + 
  ylab("Rainfall Amount") +
  theme_bw() +
  theme(#axis.title = element_blank(),
        axis.text = element_blank(),
        legend.position = c(0.3, 0.2))
# region_b
```

<!-- <a id="my_chunk"></a> -->
<!-- See [this code chunk](#my_chunk) -->

```{r riskregions, echo = FALSE,  out.width="600px", out.height="400px", fig.cap = figRef("riskregions", cap)}
plot_grid(region_a, region_b, labels = "AUTO")
cap <- "Examples of risk regions for a Rainfall-Surge compound event"
```

Of particular concern in the Netherlands are mechanisms 1 and 3, as flood risk in the canal belt system is heavily mitigated by gravity-based sluices (Figure `r figRef("sluice")`). These sluices are only operational when the internal water level is higher than the external level and the water can be discharged using gravity. For many sluices in the Netherlands, the difference between water levels needed for operation is 0.1 m, any less than this and water can flow in the opposite direction due to density differences between fresh and salt water. 

For flood risk, a core scenario for concern is therefore when excess water in the canal belt system from heavy rainfall can not be discharged due to a storm-surge raising the external water level. The important distinction here is that for this scenario the height of the storm-surge does not necessarily need to be extreme, but the level is high enough that sluice operation is not possible. The statistical problem therefore can be rephrased as a censored problem, where the occurrence of a storm surge exceeding a given level to more important than the amount it exceeds the level. This raises the natural question of how strong is the correlations between the duration of a storm surge and it's height.

```{r sluice, echo = FALSE,  out.width="600px", out.height="400px", fig.cap = figRef("sluice", cap)}
source("code/schematic_sluice.R")
sluice_plot
cap <- "Schematic of the gravity based sluice at Harlingen. Water can be discharged from the sluice provided there is a water level on the canal belt side is 0.1 m higher or more than the sea water level. The normal height of water in the canal belt system and optimal for operational purposes is -0.52 m (NAP) (Green dashed line). Operators are concerned for flooding risk when the inland water level is in excess of -0.32 m (NAP) (red dotted line)." 
```

## Event definition

In order to estimate the probability of an event occuring in one of the risk regions, we first need to define the set of variables we will use to model the problem. These variables are given in Table `r figRef("variables")`. We have adopted a similar problem specification to Li et al. 2014, who considered an application of wave heights for Dutch storms. Namely in our specification, we are interested in the duration that the gravity based is not operational. A schematic showing these variables is given in Figure `r figRef("schematics")`.

```{r variables, echo = F, tab.cap = tabRef("variables", cap)}
rv_df <- tribble(
  
  ~rv,  ~var_type,
  "T_s", "Duration of the storm surge",
  "H_s", "Maximum surge height during T_s",
  # "I", "Interarrival time between surge events",
  "H_t",  "Threshold water level that water needs to be discharged to prevent flood risk",  
  "T_d", "Duration that the gravity based sluice is not operational",
  "T_c", "Duration of common period between T_s and T_l",
  "M", "Minimum period of time the sluice needs to be operational to relieve flood risk"
  
) %>% 
  set_names(. , nm = c("Variable Name", "Variable Type"))

knitr::kable(rv_df)
```

```{r schematicSurge, echo = FALSE}
num_pi = 7
theta = seq(-pi, num_pi*pi, length.out = 360*(num_pi+ 1)/2)
tide_cycle = sin(theta)
surge = sin(1/4*theta) + 0.2
surge = surge*(surge > 0)
sea_level = tide_cycle + surge
H_t = -0.6
surge_df <- data.frame(theta, tide_cycle, sea_level)
block_x <- theta[range(which(sea_level > H_t & surge > 0))]
surge_plot <- ggplot() +
  geom_line(data = surge_df, aes(x = theta, y = tide_cycle), col = "blue", linetype = "dotted") +
  geom_line(data = surge_df, aes(x = theta, y = sea_level)) + 
  geom_hline(yintercept = H_t, linetype = "dashed") + 
  geom_polygon(data = NULL, aes(x = c(block_x, block_x[2:1]), y = c(-1, -1, H_t ,H_t)), 
            fill = "red", alpha = 0.2) +
  ylab("Water Level") +
  ylim(lims = c(-1,2.25)) +
  xlab("Time") +
  ggtitle("Example of a storm surge")
```

```{r schematicKing, echo = FALSE}
num_pi = 7
theta = seq(-pi, num_pi*pi, length.out = 360*(num_pi+ 1)/2)
shift = c(seq(0, 0.4, length.out = 200), seq(0.4, 0.4, length.out = 600),
                            seq(0.4, 0, length.out = 200),
                            rep(0, length(theta) - 1000))
tide_cycle = sin(theta) + shift + 0.1
sea_level = tide_cycle 
H_t = -0.6
close = which(abs(sea_level - H_t) < 0.02)
block_x = theta[close[diff(close) > 1][2:3]]
king_df <- data.frame(theta, tide_cycle, sea_level)
king_plot <- ggplot() +
  geom_line(data = king_df, aes(x = theta, y = tide_cycle), col = "blue", linetype = "dotted") +
  geom_line(data = king_df, aes(x = theta, y = sea_level)) + 
  geom_hline(yintercept = H_t, linetype = "dashed") + 
  geom_polygon(data = NULL, aes(x = c(block_x, block_x[2:1]), y = c(-1, -1, H_t ,H_t)), 
            fill = "red", alpha = 0.2) +
  ylab("Water Level") + 
  ylim(lims = c(-1,2.25)) +
  xlab("Time") +
  ggtitle("King tide with no surge")
```

```{r schematics, echo = FALSE,  out.width="600px", out.height="400px", fig.cap = figRef("schematics", cap)}
plot_grid(surge_plot, king_plot, labels = "AUTO")
cap <- "Red shaded areas shows the time period that gravity-based sluice was not operational was not operational for these examples. The tidal cycle is shown by the dotted blue line, and the sea level, which is the combined tide and storm surge is shown in solid black. If there is no storm surge, as in Figure 1B, then the two lines are superimposed. The dashed horizontal line shows the maximum water level for sluice operation."
```

## Navigation Links

Continue to background information about the [Tidal Data](details-on-tidal-summaries.html)  
Continue to analysis about tidal climate [Tidal Climatology](surge_climatology.html)  
