---
title: "Problem Specification"
# author: "katerobsau"
# date: "2019-08-05"
output: workflowr::wflow_html
editor_options:
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(cowplot)
library(knitr)
knitr::opts_knit$set(eval.after = 'fig.cap')
source("code/figure_labels.R")
```

<style>
p.comment {
background-color: #DBDBDB;
padding: 10px;
border: 1px solid black;
margin-left: 25px;
border-radius: 5px;
font-style: italic;
}
</style>

## Compound Events

Individually, both extreme rainfall and storm surge have the potential to produce flooding. However,compared with either event occurring individual, the potential risk is magnified if these events co-occur. This is what is known as a compound event.

<p class="comment">
**Definition 1:** The IPCC SREX (Ref 7, p. 118) defines a compound event as:  
(1) two or more extreme events occurring simultaneously or successively,   
(2) combinations of extreme events with underlying conditions that amplify the impact of the events, or   
(3) combinations of events that are not themselves extremes but lead  
to an extreme event or impact when combined.  
The contributing events can be of similar (clustered multiple events) or different type(s).
</p>

<!-- <div class="alert alert-info"> -->
<!--   Different box example -->
<!-- </div> -->

## Rainfall-Surge Mechansims

Wahl et al. (2015) breaks the compound event involving rainfall-surge down into three mechanism types. 

<p class="comment">
**Mechanism (1)** in estuarine regions, the joint occurrence of both may elevate water levels to a point where flooding is initiated or its impacts exacerbated.   
**Mechanism (2)** occurs when a destructive storm surge already causes widespread flooding, such that any significant rainfall on top of this—even if it is not an extreme event on its own—increases the flood depth and/or extent of the inundated area.   
**Mechanism (3)** occurs during a moderate storm surge which does not directly cause flooding, but is high enough to fully block or slow down gravity-fed storm water drainage, such that precipitation is more likely to cause flooding.
</p>

An example of the risk regions related to these three mechanisms is shown in Figure `r figRef("riskregions")`.

```{r region a, echo = FALSE}
delta = 0.75
regions <- data.frame(
  Risk = rep(c("M3: Rain-only", "M1: Both", "M2: Surge-only"), each = 4),
  x = c(c(0, delta, delta, 0), 
        c(delta, 1, 1, delta), 
        c(delta, 1, 1, delta)),
  y = c(c(delta, delta, 1, 1),
        c(delta, delta, 1, 1),
        c(0, 0, delta, delta))
)

region_a <- ggplot() + 
  geom_polygon(data = regions, aes(x = x, y = y, fill = Risk), alpha = 0.75) +
  geom_hline(yintercept = delta) + 
  geom_vline(xintercept = delta) + 
  xlim(c(0,1)) +
  xlab("Surge Height") + 
  ylab("Rainfall Amount") +
  ylim(c(0,1)) + 
  theme_bw() +
  theme(#axis.title = element_blank(),
        axis.text = element_blank(), 
        legend.position = c(0.3, 0.2))
# region_a
```

```{r region b, echo = FALSE}
delta = 0.8
theta = seq(0, pi/2, length.out = 90)
theta_rain = theta[theta >= pi/2*3/5]
theta_surge = theta[theta <= pi/2*2/5]
theta_both = setdiff(theta, c(theta_rain, theta_surge))

x_rain = c(delta*cos(theta_rain), 0, cos(min(theta_rain)))
y_rain = c(delta*sin(theta_rain), 1, 1)
x_surge = c(delta*cos(theta_surge), 1, 1)
y_surge = c(delta*sin(theta_surge), sin(max(theta_surge)), 0)
x_both = c(delta*cos(theta_both), cos(max(theta_both)), 1, 1)
y_both = c(delta*sin(theta_both), 1, 1, sin(min(theta_both)))
# plot(x_both, y_both)
            
regions_circle <- data.frame(
  Risk = rep(c("M3: Rain-only", "M1: Both", "M2: Surge-only"), 
             times = c(length(x_rain),length(x_both),length(x_surge))),
  x = c(x_rain, x_both, x_surge),
  y = c(y_rain, y_both, y_surge)
)

region_b <- ggplot() + 
  geom_polygon(data = regions_circle, aes(x = x, y = y, fill = Risk), alpha = 0.75) +
  geom_line(data = NULL, aes(x = delta*cos(theta), y = delta*sin(theta))) + 
  geom_line(data = NULL, aes(x = c(delta*cos(min(theta_rain)), cos(min(theta_rain))),
                             y = c(delta*sin(min(theta_rain)) , 1))) +
  geom_line(data = NULL, aes(x = c(delta*cos(max(theta_surge)), 1),
                             y = c(delta*sin(max(theta_surge)) , sin(max(theta_surge))))) +
  xlim(c(0,1)) +
  ylim(c(0,1)) + 
  xlab("Surge Height") + 
  ylab("Rainfall Amount") +
  theme_bw() +
  theme(#axis.title = element_blank(),
        axis.text = element_blank(),
        legend.position = c(0.3, 0.2))
# region_b
```

<!-- <a id="my_chunk"></a> -->
<!-- See [this code chunk](#my_chunk) -->

<!-- ```{r riskregions, echo = FALSE,  out.width="600px", out.height="400px", fig.cap = figRef("riskregions", cap)} -->
<!-- plot_grid(region_a, region_b, labels = "AUTO") -->
<!-- cap <- "Examples of risk regions for a Rainfall-Surge compound event" -->
<!-- ``` -->

## Mechanisms in the Netherlands

In the Netherlands, flood risk in the canal belt system is heavily mitigated by gravity-based sluices (Figure `r figRef("sluice")`). These sluices are only operational when the internal water level is higher than the external level and the water can be discharged using gravity. For many sluices in the Netherlands, the difference between water levels needed for operation is 0.1 m, any less than this and water can flow in the opposite direction due to density differences between fresh and salt water. 

```{r sluice, echo = FALSE,  out.width="600px", out.height="400px", fig.cap = figRef("sluice", cap)}
source("code/schematic_sluice.R")
sluice_plot
cap <- "Schematic of the gravity based sluice at Harlingen. The normal height of water in the canal belt system and optimal for operational purposes is -0.52 m (NAP) (Green dashed line). Operators are alert for flooding risk when the inland water level is in excess of -0.32 m (NAP) (red dotted line)." 
```

Given the operation of these sluices, the surge-rainfall mechanisms of 1 and 3 are of primary concern to the Netherlands. In these instances, the sluice is not operational. Figure `r figRef("schematics")` shows two simple examples where the external water level prevents the release of water; one due to a surge event and the other due to a high tide.

```{r schematicSurge, echo = FALSE}
num_pi = 7
theta = seq(0, num_pi*pi, length.out = 360*(num_pi+ 1)/2)
tide_cycle = sin(theta)
surge = sin(1/4*theta) + 0.35
surge = surge*(surge > 0)
sea_level = tide_cycle + surge
H_t = -0.42
surge_df <- data.frame(theta, tide_cycle, sea_level, surge)
clusters = rle(sea_level > H_t)
i = 1:(clusters$length[1])
block_x_sur <- range(theta[i])
surge_plot <- ggplot() +
  geom_line(data = surge_df, aes(x = theta, y = surge), col = "red", linetype = "dotted") +
  geom_line(data = surge_df, aes(x = theta, y = tide_cycle), col = "blue", linetype = "dashed") +
  geom_line(data = surge_df, aes(x = theta, y = sea_level)) + 
  geom_hline(yintercept = H_t, linetype = "dashed") + 
  geom_polygon(data = NULL, aes(x = c(block_x_sur, block_x_sur[2:1]), y = c(-1, -1, H_t ,H_t)), 
            fill = "red", alpha = 0.2) +
  ylab("Water Level") +
  ylim(lims = c(-1,2.25)) +
  xlab("Time") +
  ggtitle("Example of a storm surge") + 
  theme_bw()
```

```{r schematicKing, echo = FALSE}
num_pi = 7
theta = seq(0, num_pi*pi, length.out = 360*(num_pi+ 1)/2)
shift = c(seq(0, 0.4, length.out = 200), seq(0.4, 0.4, length.out = 600),
                            seq(0.4, 0, length.out = 200),
                            rep(0, length(theta) - 1000))
tide_cycle = sin(theta) + shift + 0.2
sea_level = tide_cycle 
surge = rep(0, length(sea_level)) 
H_t = -0.42
close = which(abs(sea_level - H_t) < 0.02)
clusters = rle(sea_level > H_t)
i = 1:(clusters$length[1])
block_x <- range(theta[i])
king_df <- data.frame(theta, tide_cycle, sea_level, surge)
king_plot <- ggplot() +
  geom_line(data = king_df, aes(x = theta, y = tide_cycle), col = "blue", linetype = "dashed") +
  geom_line(data = king_df, aes(x = theta, y = surge), col = "red", linetype = "dotted") +
  geom_line(data = king_df, aes(x = theta, y = sea_level)) + 
  geom_hline(yintercept = H_t, linetype = "dashed") + 
  geom_polygon(data = NULL, aes(x = c(block_x, block_x[2:1]), y = c(-1, -1, H_t ,H_t)), 
            fill = "red", alpha = 0.2) +
  ylab("Water Level") + 
  ylim(lims = c(-1,2.25)) +
  xlab("Time") +
  ggtitle("King tide with no surge") + 
  theme_bw()
```

```{r schematics, echo = FALSE,  out.width="600px", out.height="400px", fig.cap = figRef("schematics", cap)}
plot_grid(surge_plot, king_plot, labels = "AUTO")
cap <- "Two examples showing when the sea level (solid back) exceeds the flood risk level (dashed black). The red shaded areas show the period when the gravity-based sluice was not operational. Here the tidal cycle is shown by the dashed blue line, and the surge in the red dotted line. The tide and surge combined give the total level. If there is no storm surge, as in Figure 1B, then the two lines are superimposed. "
```

## Rethinking the Risk Framework 

An important distinction when assessing the risk of rainfall-surge events is that in the Netherlands the height of the storm-surge does not necessarily need to be extreme. Instead, of concern is 

* Does the total sea level (tide + surge) exceed the risk threshold?

* How long does the total level stay above this risk threshold, preventing discharge from the sluice?

In this sense the statistical problem becomes a censored one, as we are not concerned with the extremeness of the surge height. Once the surge contributes to the total level exceeding the risk threshold, it doesn't matter by how much the threshold is exceeded, the sluice is not operational. This requires a fundamental reframing of the statistical problem, and move away from the traditional statistical approach of assessing risk by modelling the dependence between storm surge heights and rainfall amounts.

The reframing of the problem may also be advantageous. In misspecifying the problem, risk is not reliably assessed. Additionally, the dependence between extreme rainfall and extreme surge heights is asymptotically weak at many locations throughout the world, including the Netherlands. (cite Wahl, Westra, Bevacqua). This weak dependence is difficult to accurately estimate (Westra) and as a consequence it may even be beneficial to model the events independently despite the compound nature of their impacts. A rephrasing of the problem may be beneficial to improving estimates of dependence parameters and quantifying risk. 

## Navigation Links

* [Problem Specification](compound-event-definition.html)  
* [Background to Tidal Data](details-on-tidal-summaries.html)
* [Tidal Climatology](surge_climatology.html)  
* [Storm Surge Ensemble](ensemble_details.html)  
* [Univariate Post-Processing](univariate_pp.html)  
* [Event Based Verification](event-based-verification.html)
