---
title: "Surge Climatology"
output: html_document
---

```{r loadPackages, echo = FALSE, warning = FALSE, message = FALSE}
library(tidyverse)
library(cowplot)
source("code/figure_labels.R")
source("code/source/surge_exploratory_analyis.R")
```

```{r tidalPackage read, echo = FALSE, warning = FALSE, message = FALSE}
package_dir = "code/tidalHelpers"
install.packages(pkgs = "code/tidalHelpers", repos = NULL, type = "source")
library(tidalHelpers)
```

In this section, the available observational data is visualised and the conditions when the sluice is not operational are examined.

## Harlingen Observations

The following table displays the observed data at the Harlingen station. The storm surge is obtained by subtracting the harmonic tide from the observed sea level, see Figure `r figRef("plotSurgeObs")`. 

```{r tableObs, echo = FALSE}
obs_data %>%
  setNames(nm = c("Date", "Harmonic Tide", "Observation", "Surge")) %>%
  head() %>%
  knitr::kable()
```

```{r plotObs, echo = FALSE, fig.cap = figRef("plotSurgeObs", cap)}
len = 5*25*60/10
ggplot(obs_data[1:len,]) +
  geom_path(aes(x = ldate, y = harm), col = "blue", linetype = "dashed") + 
  geom_path(aes(x = ldate, y = obs), col = "black") +
  geom_path(aes(x = ldate, y = sur), col = "red", linetype = "dotted", size = 1.25) +
  geom_hline(yintercept = 0, linetype = "dashed", col = "grey") +
  # geom_hline(yintercept = -0.5, linetype = "dashed") +
  ylab("Water Level") +
  xlab("Date") + 
  theme_bw()
cap <- "Plot showing the sea level (solid black), harmonic tide (blue dashed) and strom surge (red dotted) observations  Example"
```

## Summary Information

```{r sluiceHeight, echo = FALSE}
#h_t_level = -0.42 
# included in exploratory analysis
obs_vec = obs_data %>% pull(obs) 
h_t_quantile = sum(obs_vec <= h_t_level, na.rm = TRUE)/sum(!is.na(obs_vec))
```

Summary of Harlingen observational data:

* The date range of the available observations is `r range(obs_data$ldate)[1]` to `r range(obs_data$ldate)[2]` .
* There are two years of missing data (will visualise at a later date).
* The tidal range is `r range(obs_data$harm)` m. 
* The surge range is `r range(obs_data$sur, na.rm = TRUE)` m. 
* Largest observated sea level was recorded on `r obs_data$ldate[which.max(obs_data$obs)]` of `r obs_data$obs[which.max(obs_data$obs)]` m corresponding to a `r obs_data$sur[which.max(obs_data$obs)]` m surge.
* Lowest observated sea level was recorded on `r obs_data$ldate[which.min(obs_data$obs)]` of `r obs_data$obs[which.min(obs_data$obs)]` m corresponding to a `r obs_data$sur[which.min(obs_data$obs)]` m surge.
* A threshold height of `r h_t_level` corresponds to a quantile of `r h_t_quantile` in observed water levels.

## Sluice Operation 

For a sluice height of `r h_t_level` m based on the historical observation, the histogram shows the number of times and for how long the sluices was not operational.

<!-- ```{r getClusters, echo = FALSE} -->
<!-- flood_risk = obs_data %>% -->
<!--   mutate(bool_col = obs >= h_t_level) -->

<!-- clusters = flood_risk %>% -->
<!--   pull(bool_col) %>% -->
<!--   rle() %>% -->
<!--   unclass() %>% -->
<!--   as.data.frame() %>% -->
<!--   mutate(end = cumsum(lengths), -->
<!--        start = c(1, dplyr::lag(end)[-1] + 1)) %>% -->
<!--   mutate(date_end = obs_data$ldate[end]) %>% -->
<!--   mutate(date_start = obs_data$ldate[start]) %>% -->
<!--   # select(-start, -end) %>% -->
<!--   arrange(desc(lengths)) %>% -->
<!--   mutate(temp = paste(0, lengths*10, 0, sep = ":")) %>% -->
<!--   mutate(time = lubridate::hms(temp, roll = TRUE)) %>% -->
<!--   select(-temp)  -->

<!-- print("Need to specify a minimum interarrival time for discharge water") -->
<!-- print("Need to check how rle() handles missing data") -->
<!-- ``` -->

```{r displayClusters, echo = FALSE}
clusters_risk_threshold %>%
  # filter(values == TRUE) %>%
  select(-values, -lengths) %>%
  select(time, date_start, date_end) %>%
  head() %>%
  knitr::kable(align = "l")
```

```{r histPlot, echo = FALSE}
plot_data <- clusters_risk_threshold %>% 
  # dplyr::filter(values == TRUE) %>% 
  mutate(lengths = lengths*10/60/24)
med_len = median(plot_data %>% pull(lengths))
mean_len = mean(plot_data %>% pull(lengths))
mean_hms = paste(0, floor(mean_len*24*60), 0, sep = ":") %>% lubridate::hms(roll = TRUE) %>% as.character()
hist_plot <- ggplot(plot_data) + 
  geom_histogram(aes(x = lengths), bins = 100) +
  geom_vline(xintercept = mean_len, 
             col = "red", linetype = "dotted") +
  theme_bw() + 
  xlab("Days")
hist_plot

clusters = clusters_risk_threshold

```

There are `r nrow(plot_data)` times that the sluice is not operational, and the average period that the sluice is not in operational is `r mean_hms`. This is intuitive given the average period spent in the high tide cycle. Currently we do not condition on the minimum period the sluice can be operated.

## Longest Periods above the Threshold

The four longest periods when the sluice was not operational are shown in the below figure. We observe in Figure B there is a problem with the underlying data, however during this period the sluice was operational due to surge as the plot indicates.

```{r sluiceExamples, echo = FALSE}
# print("Note: Wrap repition of plots in a function")
lens = clusters$lengths[1:4]
num_days = (10*24)*60/10 

# delta is used to make the x axis size the same across the four plots
delta1 = lubridate::minutes(max((num_days - lens[1])/2, 0)*10)
p1 <- ggplot(obs_data %>% filter(
    ldate >= (clusters$date_start[1] - delta1) & 
    ldate <= (clusters$date_end[1] + delta1))) + 
  geom_path(aes(x = ldate, y = harm), col = "blue", linetype = "dashed", na.rm = TRUE) + 
  geom_path(aes(x = ldate, y = obs), col = "black", na.rm = TRUE) +
  geom_path(aes(x = ldate, y = sur), col = "red", linetype = "dotted", size = 1.25, na.rm = TRUE) +
  geom_hline(yintercept = -0.5, linetype = "dashed") +
  ylab("Water Level") +
  xlab("Date") + 
  ylim(c(-1.5, 3.5)) +
  theme_bw()

delta2 = lubridate::minutes(max((num_days - lens[2])/2, 0)*10)
p2 <- ggplot(obs_data %>% filter(
    ldate >= (clusters$date_start[2] - delta2) & 
    ldate <= (clusters$date_end[2] + delta2)   )) + 
  geom_path(aes(x = ldate, y = harm), col = "blue", linetype = "dashed", na.rm = TRUE) + 
  geom_path(aes(x = ldate, y = obs), col = "black", na.rm = TRUE) +
  geom_path(aes(x = ldate, y = sur), col = "red", linetype = "dotted", size = 1.25, na.rm = TRUE) +
  geom_hline(yintercept = -0.5, linetype = "dashed") +
  ylab("Water Level") +
  xlab("Date") + 
  ylim(c(-1.5, 3.5)) +
  theme_bw()

delta3 = lubridate::minutes(max((num_days - lens[3])/2, 0)*10)
p3 <- ggplot(obs_data %>% filter(ldate >= (clusters$date_start[3] - delta3) & 
                          ldate <= (clusters$date_end[3] + delta3)   )) + 
  geom_path(aes(x = ldate, y = harm), col = "blue", linetype = "dashed") + 
  geom_path(aes(x = ldate, y = obs), col = "black") +
  geom_path(aes(x = ldate, y = sur), col = "red", linetype = "dotted", size = 1.25) +
  geom_hline(yintercept = -0.5, linetype = "dashed") +
  ylab("Water Level") +
  xlab("Date") + 
  ylim(c(-1.5, 3.5)) +
  theme_bw()

delta4 = lubridate::minutes(max((num_days - lens[4])/2, 0)*10)
p4 <- ggplot(obs_data %>% filter(ldate >= (clusters$date_start[4] - delta4) & 
                          ldate <= (clusters$date_end[4] + delta4))) + 
  geom_path(aes(x = ldate, y = harm), col = "blue", linetype = "dashed") + 
  geom_path(aes(x = ldate, y = obs), col = "black") +
  geom_path(aes(x = ldate, y = sur), col = "red", linetype = "dotted", size = 1.25) +
  geom_hline(yintercept = -0.5, linetype = "dashed") +
  ylab("Water Level") +
  xlab("Date") + 
  ylim(c(-1.5, 3.5)) +
  theme_bw()

plot_grid(p1, p2, p3, p4, labels = "AUTO", ncol = 2)
```

## Highest Surges on Record

The four highest surges are shown in the below Figure. There is not a one-to-one correspondance with the highest surge events and the longest duration events where the sluice was unable to be operated to prevent flood risk. 

```{r highestSurges, echo = FALSE}
delta_day = num_days/2
num = 4
bool_val = TRUE
min_period = 5*24*60/10
while(bool_val == TRUE){
  max_surge_i = order(obs_data$sur, decreasing = TRUE)[1:num]
  diff_i = sort(obs_data$ldate[max_surge_i]) %>% diff()
  bool_val = sum(diff_i > min_period) != 3
  num = num + 1
  # print(num)
}

temp = sort(obs_data$ldate[max_surge_i])[c(which(diff_i >= min_period), length(max_surge_i))]
max_surge_i = which(obs_data$ldate %in% temp)

s1 <- ggplot(obs_data[(max_surge_i[1] - delta_day):(max_surge_i[1] + delta_day), ]) + 
  geom_path(aes(x = ldate, y = harm), col = "blue", linetype = "dashed") + 
  geom_path(aes(x = ldate, y = obs), col = "black") +
  geom_path(aes(x = ldate, y = sur), col = "red", linetype = "dotted", size = 1.25) +
  geom_hline(yintercept = -0.5, linetype = "dashed") +
  ylab("Water Level") +
  xlab("Date") + 
  ylim(c(-1.5, 3.5)) +
  theme_bw()

s2 <- ggplot(obs_data[(max_surge_i[2] - delta_day):(max_surge_i[2] + delta_day), ]) + 
  geom_path(aes(x = ldate, y = harm), col = "blue", linetype = "dashed") + 
  geom_path(aes(x = ldate, y = obs), col = "black") +
  geom_path(aes(x = ldate, y = sur), col = "red", linetype = "dotted", size = 1.25) +
  geom_hline(yintercept = -0.5, linetype = "dashed") +
  ylab("Water Level") +
  xlab("Date") + 
  ylim(c(-1.5, 3.5)) +
  theme_bw()

s3 <- ggplot(obs_data[(max_surge_i[3] - delta_day):(max_surge_i[3] + delta_day), ]) + 
  geom_path(aes(x = ldate, y = harm), col = "blue", linetype = "dashed") + 
  geom_path(aes(x = ldate, y = obs), col = "black") +
  geom_path(aes(x = ldate, y = sur), col = "red", linetype = "dotted", size = 1.25) +
  geom_hline(yintercept = -0.5, linetype = "dashed") +
  ylab("Water Level") +
  xlab("Date") + 
  ylim(c(-1.5, 3.5)) +
  theme_bw()

s4 <- ggplot(obs_data[(max_surge_i[4] - delta_day):(max_surge_i[4] + delta_day), ]) + 
  geom_path(aes(x = ldate, y = harm), col = "blue", linetype = "dashed") + 
  geom_path(aes(x = ldate, y = obs), col = "black") +
  geom_path(aes(x = ldate, y = sur), col = "red", linetype = "dotted", size = 1.25) +
  geom_hline(yintercept = -0.5, linetype = "dashed") +
  ylab("Water Level") +
  xlab("Date") + 
  ylim(c(-1.5, 3.5)) +
  theme_bw()

plot_grid(s1, s2, s3, s4, labels = "AUTO", ncol = 2)
```

## Relationship between Duration and Surge 

```{r echo = FALSE, warning = FALSE}
# max_height = apply(clusters, 1, function(r, flood_risk, col_names){
#   i1 = r[which(col_names == "start")]
#   i2 = r[which(col_names == "end")]
#   flood_risk[i1:i2,] %>%
#     select(harm, obs, sur) %>%
#     apply(., 2, max)
# }, flood_risk, col_names = names(clusters)) 
# max_height = t(max_height)

plot_clusters <- cbind(clusters_risk_threshold, max_height) 

duration_vs_sur<- ggplot(plot_clusters) + 
  geom_point(aes(x = time@hour, y = sur)) + #time@hour
  # geom_smooth(aes(x = time@hour, y = sur)) +
  geom_vline(xintercept = seq(0, 120, 6),
             col = "red", linetype = "dotted") +
  theme_bw() + 
  xlab("Duration external water level exceeded the risk level") + 
  ylab("Surge Height") 
duration_vs_sur
```

As the period that the external water level exceeds `r h_t_level` m increases, there is not a strong correlation with an increase in the surge height. This suggests that the important factor for flood mitigation in the instance of the sluice operation, is not infact surge height, but the time period spent above the risk threshold. This is in contrast to the many frameworks who consider the problem of compound events from the perspective of heavy rainfall and high surge combination.

We note that although the duration is a continuous random variable, the duration values appear binned due to the underlying influence of the tidal cycle. This is because the period above the risk threshold often corresponds with the rising tide and the end of the risk period with the outgoing tide with some additional variability due to the surge. The difference between these cycles is roughly a 12 hours. 

## Surge duration

To analyse a random variable that is a bit friendlier in terms of dependence and standard copulas, it is useful to think about how the duration of the surge event is related to the surge height. The points are shaded by the percentage of time spent about the flood risk threshold during each storm surge event.

```{r clustersSur, echo = FALSE}
# clusters_sur = flood_risk %>%
#   mutate(bool_sur = sur > 0) %>%
#   pull(bool_sur) %>%
#   rle() %>%
#   unclass() %>%
#   as.data.frame() %>%
#   mutate(end = cumsum(lengths),
#        start = c(1, dplyr::lag(end)[-1] + 1)) %>%
#   mutate(date_end = obs_data$ldate[end]) %>%
#   mutate(date_start = obs_data$ldate[start]) %>%
#   # select(-start, -end) %>%
#   arrange(desc(lengths)) %>%
#   mutate(temp = paste(0, lengths*10, 0, sep = ":")) %>%
#   mutate(time = lubridate::hms(temp, roll = TRUE)) %>%
#   select(-temp) 
# 
# max_height_sur = apply(clusters_sur, 1, function(r, flood_risk, col_names){
#   i1 = r[which(col_names == "start")]
#   i2 = r[which(col_names == "end")]
#   flood_risk[i1:i2,] %>%
#     select(harm, obs, sur) %>%
#     apply(., 2, max)
# }, flood_risk, col_names = names(clusters_sur)) 
# max_height_sur = t(max_height_sur)
# 
# thres_exceeded = apply(clusters_sur, 1, function(r, flood_risk, col_names){
#   i1 = r[which(col_names == "start")] %>% as.numeric()
#   i2 = r[which(col_names == "end")] %>% as.numeric()
#   temp = flood_risk[i1:i2,] %>%
#     pull(obs)
#   perc = sum(temp > h_t_level)/sum(!is.na(temp))
# }, flood_risk, col_names = names(clusters_sur)) 

plot_clusters_sur <- cbind(clusters_sur, max_height_sur) %>%
  mutate(risk = thres_exceeded %>% as.numeric())

duration_vs_sur1 <- ggplot(plot_clusters_sur %>% filter(sur > 0.01)) + 
  geom_point(aes(x = time@hour, y = sur, col = risk), size = 0.75) + #time@hour
  scale_color_distiller("Percentage", palette = "RdBu") +
  theme_bw() + 
  xlab("Surge Duration") + 
  ylab("Surge Height") 

duration_vs_sur1
```

<!-- ```{r} -->
<!-- duration_vs_sur2 <- ggplot(plot_clusters_sur %>% filter(sur > 0.01)) +  -->
<!--   geom_point(aes(x = time@hour*risk, y = sur), size = 0.75) + #time@hour -->
<!--   scale_color_distiller("Percentage", palette = "RdBu") + -->
<!--   theme_bw() +  -->
<!--   xlab("Duration above the risk threshold during the Storm Surge") +  -->
<!--   ylab("Surge Height")  -->

<!-- duration_vs_sur2 -->
<!-- ``` -->


<!-- ## To Do: -->

<!-- Ideally also like to wrap plotting code, and to fix dates on axes. -->

## Navigation Links

* [Project Overview](index.html)  
* [Sea Level Metadata and Preprocessing](details-on-tidal-summaries.html)  
* [Rainfall Metadata and Preprocessing ](details-on-rainfall-data.html)  
* [Surge Climatology](surge_climatology.html)  
* [Surge Seasonality](surge-seasonality.html)  
