---
title: "Surge Climatology"
output: html_document
---

```{r loadPackages, echo = FALSE, warning = FALSE, message = FALSE}
library(tidyverse)
library(cowplot)
source("code/figure_labels.R")
```

```{r tidalPackage read, echo = FALSE, warning = FALSE, message = FALSE}
package_dir = "code/tidalHelpers"
install.packages(pkgs = "code/tidalHelpers", repos = NULL, type = "source")
library(tidalHelpers)
```

**WORK IN PROGRESS**

In this section, the available observational data is visualised and the conditions when the sluice is not operational are examined.

## Observations

The following table displays the observed data at the Harlingen station. The storm surge is obtained by subtracting the harmonic tide from the observed sea level, see Figure `r figRef("plotSurgeObs")`. 

```{r readObs, echo = FALSE, warning = FALSE, message = FALSE}
ens_summary_file = "Ensemble_Summary.rds"
init_var = tidalHelpers::utils_init()
ens_summary_dir = init_var$save_dir
ens_summary_path = paste(ens_summary_dir, ens_summary_file, sep = "")
ens_summary = readRDS(ens_summary_path) 

obs_data = ens_summary %>% 
  dplyr::filter(t < 24) %>%
  dplyr::mutate(
    temp_date = gsub("-", " ", date) %>% lubridate::ymd()) %>%
  dplyr::mutate(
    temp_time = paste(0, round(t*60), 0, sep = " ") %>% lubridate::hms(roll = TRUE)) %>%
  dplyr::mutate(ldate = 
    lubridate::ymd_hms(temp_date + temp_time)) %>%
  dplyr::select(ldate, harm, obs, sur) %>%
  dplyr::mutate(obs = if_else(is.na(sur), NA_real_, obs))

print("Note: Should add handling of NA Surge values into preprocessing")
```

```{r tableObs, echo = FALSE}
obs_data %>%
  setNames(nm = c("Date", "Harmonic Tide", "Observation", "Surge")) %>%
  head() %>%
  knitr::kable()

```

```{r plotObs, echo = FALSE, fig.cap = figRef("plotSurgeObs", cap)}
len = 5*25*60/10
ggplot(obs_data[1:len,]) +
  geom_path(aes(x = ldate, y = harm), col = "blue", linetype = "dashed") + 
  geom_path(aes(x = ldate, y = obs), col = "black") +
  geom_path(aes(x = ldate, y = sur), col = "red", linetype = "dotted", size = 1.25) +
  geom_hline(yintercept = 0, linetype = "dashed", col = "grey") +
  # geom_hline(yintercept = -0.5, linetype = "dashed") +
  ylab("Water Level") +
  xlab("Date") + 
  theme_bw()
cap <- "Plot showing the sea level (solid black), harmonic tide (blue dashed) and strom surge (red dotted) observations  Example"
```

## Summary Information

```{r sluiceHeight, echo = FALSE}
h_t_level = -0.42
obs_vec = obs_data %>% pull(obs) 
h_t_quantile = sum(obs_vec <= h_t_level, na.rm = TRUE)/sum(!is.na(obs_vec))
```

Summary of Harlingen observational data:

* The date range of the available observations is `r range(obs_data$ldate)[1]` to `r range(obs_data$ldate)[2]` .
* There are two years of missing data (will visualise at a later date).
* The tidal range is `r range(obs_data$harm)` m. 
* The surge range is `r range(obs_data$sur, na.rm = TRUE)` m. 
* Largest observated sea level was recorded on `r obs_data$ldate[which.max(obs_data$obs)]` of `r obs_data$obs[which.max(obs_data$obs)]` m corresponding to a `r obs_data$sur[which.max(obs_data$obs)]` m surge.
* Lowest observated sea level was recorded on `r obs_data$ldate[which.min(obs_data$obs)]` of `r obs_data$obs[which.min(obs_data$obs)]` m corresponding to a `r obs_data$sur[which.min(obs_data$obs)]` m surge.
* A threshold height of `r h_t_level` corresponds to a quantile of `r h_t_quantile` in observed water levels.

## Sluice Operation 

For a sluice height of `r h_t_level` m based on the historical observation, the histogram shows the number of times and for how long the sluices was not operational.

```{r getClusters, echo = FALSE}
flood_risk = obs_data %>%
  mutate(bool_col = obs >= h_t_level)
clusters = flood_risk %>%
  pull(bool_col) %>%
  rle() %>%
  unclass() %>%
  as.data.frame() %>%
  mutate(end = cumsum(lengths),
       start = c(1, dplyr::lag(end)[-1] + 1)) %>%
  mutate(date_end = obs_data$ldate[end]) %>%
  mutate(date_start = obs_data$ldate[start]) %>%
  # select(-start, -end) %>%
  arrange(desc(lengths)) %>%
  mutate(temp = paste(0, lengths*10, 0, sep = ":")) %>%
  mutate(time = lubridate::hms(temp, roll = TRUE)) %>%
  select(-temp) 

print("Need to specify a minimum interarrival time for discharge water")
print("Need to check how rle() handles missing data")
```

```{r displayClusters, echo = FALSE}
clusters %>%
  filter(values == TRUE) %>%
  select(-values, -lengths) %>%
  select(time, date_start, date_end) %>%
  head() %>%
  knitr::kable(align = "l")
```

```{r histPlot, echo = FALSE}
plot_data <- clusters %>% 
  dplyr::filter(values == TRUE) %>% 
  mutate(lengths = lengths*10/60/24)
med_len = median(plot_data %>% pull(lengths))
mean_len = mean(plot_data %>% pull(lengths))
mean_hms = paste(0, floor(mean_len*24*60), 0, sep = ":") %>% lubridate::hms(roll = TRUE) %>% as.character()
hist_plot <- ggplot(plot_data) + 
  geom_histogram(aes(x = lengths), bins = 100) +
  geom_vline(xintercept = mean_len, 
             col = "red", linetype = "dotted") +
  theme_bw() + 
  xlab("Days")
hist_plot
```

There are `r nrow(plot_data)` times that the sluice is not operational, and the average period that the sluice is not in operational is `r mean_hms`. This is intuitive given the average period spent in the high tide cycle. Currently we do not condition on the minimum period the sluice can be operated.

## Longest Periods above the Threshold

The four longest periods when the sluice was not operational are shown in the below figure. We observe in Figure B there is a problem with the underlying data.

```{r sluiceExamples, echo = FALSE}
print("Note: Wrap repition of plots in a function")
lens = clusters$lengths[1:4]
num_days = (10*24)*60/10 

# delta is used to make the x axis size the same across the four plots
delta1 = lubridate::minutes(max((num_days - lens[1])/2, 0)*10)
p1 <- ggplot(obs_data %>% filter(
    ldate >= (clusters$date_start[1] - delta1) & 
    ldate <= (clusters$date_end[1] + delta1))) + 
  geom_path(aes(x = ldate, y = harm), col = "blue", linetype = "dashed", na.rm = TRUE) + 
  geom_path(aes(x = ldate, y = obs), col = "black", na.rm = TRUE) +
  geom_path(aes(x = ldate, y = sur), col = "red", linetype = "dotted", size = 1.25, na.rm = TRUE) +
  geom_hline(yintercept = -0.5, linetype = "dashed") +
  ylab("Water Level") +
  xlab("Date") + 
  ylim(c(-1.5, 3.5)) +
  theme_bw()

delta2 = lubridate::minutes(max((num_days - lens[2])/2, 0)*10)
p2 <- ggplot(obs_data %>% filter(
    ldate >= (clusters$date_start[2] - delta2) & 
    ldate <= (clusters$date_end[2] + delta2)   )) + 
  geom_path(aes(x = ldate, y = harm), col = "blue", linetype = "dashed", na.rm = TRUE) + 
  geom_path(aes(x = ldate, y = obs), col = "black", na.rm = TRUE) +
  geom_path(aes(x = ldate, y = sur), col = "red", linetype = "dotted", size = 1.25, na.rm = TRUE) +
  geom_hline(yintercept = -0.5, linetype = "dashed") +
  ylab("Water Level") +
  xlab("Date") + 
  ylim(c(-1.5, 3.5)) +
  theme_bw()

delta3 = lubridate::minutes(max((num_days - lens[3])/2, 0)*10)
p3 <- ggplot(obs_data %>% filter(ldate >= (clusters$date_start[3] - delta3) & 
                          ldate <= (clusters$date_end[3] + delta3)   )) + 
  geom_path(aes(x = ldate, y = harm), col = "blue", linetype = "dashed") + 
  geom_path(aes(x = ldate, y = obs), col = "black") +
  geom_path(aes(x = ldate, y = sur), col = "red", linetype = "dotted", size = 1.25) +
  geom_hline(yintercept = -0.5, linetype = "dashed") +
  ylab("Water Level") +
  xlab("Date") + 
  ylim(c(-1.5, 3.5)) +
  theme_bw()

delta4 = lubridate::minutes(max((num_days - lens[4])/2, 0)*10)
p4 <- ggplot(obs_data %>% filter(ldate >= (clusters$date_start[4] - delta4) & 
                          ldate <= (clusters$date_end[4] + delta4))) + 
  geom_path(aes(x = ldate, y = harm), col = "blue", linetype = "dashed") + 
  geom_path(aes(x = ldate, y = obs), col = "black") +
  geom_path(aes(x = ldate, y = sur), col = "red", linetype = "dotted", size = 1.25) +
  geom_hline(yintercept = -0.5, linetype = "dashed") +
  ylab("Water Level") +
  xlab("Date") + 
  ylim(c(-1.5, 3.5)) +
  theme_bw()

plot_grid(p1, p2, p3, p4, labels = "AUTO", ncol = 2)
```

## Highest Surges on Record

The four highest surges are shown in the below Figure. There is not a one-to-one correspondance with the highest surge events and the longest events that pose flood risk. 

```{r highestSurges, echo = FALSE}
delta_day = num_days/2
num = 4
bool_val = TRUE
min_period = 5*24*60/10
while(bool_val == TRUE){
  max_surge_i = order(obs_data$sur, decreasing = TRUE)[1:num]
  diff_i = sort(obs_data$ldate[max_surge_i]) %>% diff()
  bool_val = sum(diff_i > min_period) != 3
  num = num + 1
  # print(num)
}

temp = sort(obs_data$ldate[max_surge_i])[c(which(diff_i >= min_period), length(max_surge_i))]
max_surge_i = which(obs_data$ldate %in% temp)

s1 <- ggplot(obs_data[(max_surge_i[1] - delta_day):(max_surge_i[1] + delta_day), ]) + 
  geom_path(aes(x = ldate, y = harm), col = "blue", linetype = "dashed") + 
  geom_path(aes(x = ldate, y = obs), col = "black") +
  geom_path(aes(x = ldate, y = sur), col = "red", linetype = "dotted", size = 1.25) +
  geom_hline(yintercept = -0.5, linetype = "dashed") +
  ylab("Water Level") +
  xlab("Date") + 
  ylim(c(-1.5, 3.5)) +
  theme_bw()

s2 <- ggplot(obs_data[(max_surge_i[2] - delta_day):(max_surge_i[2] + delta_day), ]) + 
  geom_path(aes(x = ldate, y = harm), col = "blue", linetype = "dashed") + 
  geom_path(aes(x = ldate, y = obs), col = "black") +
  geom_path(aes(x = ldate, y = sur), col = "red", linetype = "dotted", size = 1.25) +
  geom_hline(yintercept = -0.5, linetype = "dashed") +
  ylab("Water Level") +
  xlab("Date") + 
  ylim(c(-1.5, 3.5)) +
  theme_bw()

s3 <- ggplot(obs_data[(max_surge_i[3] - delta_day):(max_surge_i[3] + delta_day), ]) + 
  geom_path(aes(x = ldate, y = harm), col = "blue", linetype = "dashed") + 
  geom_path(aes(x = ldate, y = obs), col = "black") +
  geom_path(aes(x = ldate, y = sur), col = "red", linetype = "dotted", size = 1.25) +
  geom_hline(yintercept = -0.5, linetype = "dashed") +
  ylab("Water Level") +
  xlab("Date") + 
  ylim(c(-1.5, 3.5)) +
  theme_bw()

s4 <- ggplot(obs_data[(max_surge_i[4] - delta_day):(max_surge_i[4] + delta_day), ]) + 
  geom_path(aes(x = ldate, y = harm), col = "blue", linetype = "dashed") + 
  geom_path(aes(x = ldate, y = obs), col = "black") +
  geom_path(aes(x = ldate, y = sur), col = "red", linetype = "dotted", size = 1.25) +
  geom_hline(yintercept = -0.5, linetype = "dashed") +
  ylab("Water Level") +
  xlab("Date") + 
  ylim(c(-1.5, 3.5)) +
  theme_bw()

plot_grid(s1, s2, s3, s4, labels = "AUTO", ncol = 2)
```

## Relationship between Duration and Surge 

```{r echo = FALSE}
max_height = apply(clusters, 1, function(r, flood_risk, col_names){
  i1 = r[which(col_names == "start")]
  i2 = r[which(col_names == "end")]
  flood_risk[i1:i2,] %>%
    select(harm, obs, sur) %>%
    apply(., 2, max)
}, flood_risk, col_names = names(clusters)) 
max_height = t(max_height)

plot_clusters <- cbind(clusters, max_height) 

duration_vs_sur<- ggplot(plot_clusters) + 
  geom_point(aes(x = time@hour, y = sur)) + #time@hour
  geom_vline(xintercept = seq(0, 120, 6),
             col = "red", linetype = "dotted") +
  theme_bw() + 
  xlab("Time in hours") + 
  ylab("Surge Height") 
duration_vs_sur
```

As the period that the external water level exceeds `r h_t_level` m increases, there is not a strong correlation with an increase in the surge height. This suggests that the important factor for flood mitigation in the instance of the sluice operation, is not infact surge height, but the time period spent above the risk threshold. This is in contrast to the many frameworks who consider the problem of compound events from the perspective of heavy rainfall and high surge combination.

We note that the duration is a continuous random variable, but that due to the underlying influence of the tidal cycle, that the duration values appear binned. This is because the risk period above the threshold often corresponds with the rising tide and the end of the risk period with the outgoing tide. The difference between these cycles is roughly a factor of 12 hours. 

<!-- ## To Do: -->

<!-- Ideally also like to wrap plotting code, and to fix dates on axes. -->
