---
title: "Rainfall: Univariate Post-processing"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(tidyverse)
library(plotly)
library(MASS)
```

## Univariate Model 

It is common to observe days without any rainfall. Therefore when modelling rainfall, the distribution used needs to account for the presence of zero-inflated observations, such as the Zero adjusted gamma distribution (ZAGA). Within the Netherlands, this distribution has been used for the statistical post-processing of rainfall, with improved predictive performance demonstrated over the raw ensemble (Whan and Schmeits 2018). Possible alternative distributions include the Censored Shifted Guassion distribution (CSG) (Scheuerer et al. 2017), and the lognormal or inverse Gaussian distributions (Bentzien and Friederichs 2012). However, as we are primarily interested in bivariate processing, we use the ZAGA distribution as our baseline and leave comparisons with other univariate distributions for a different study.

The Zero adjusted Gamma distribution is 

$$ f(y \mid \mu, \sigma, \nu) = 
\begin{cases} 
    \nu, & \text{for } y = 0 \\
    (1 - \nu) g(y \mid  \mu, \sigma, \nu) , & \text{for } y > 0, 
\end{cases}$$

where $g(\cdot)$ is the gamma distribution

$$g(y \mid \mu,\sigma,\nu) =   \dfrac{y^{1/ \sigma^2 - 1} \exp\left(-y  / (\sigma^2\mu) \right) }{(\mu\sigma^2)^{1/ \sigma^2} \Gamma(1/\sigma^2)}$$,

with support $y \in (0,\infty)$, and parameters $\mu>0$, $\sigma>0$ and $0<\nu<1$.

Parameters of this distribution are fitted using a generalised additive model for distibutions in the location, scale and shape family (gamlss) (R. A. Rigby and D. M. Stasinopoulos et al. 2007). 
The parameters of this distribution are modelled using covariates estimated from the raw ensemble, which include the ensemble mean, ensemble standard deviation and an empirically estimated $\nu$. Inclusion of these covariates in the gamlss model is determined using forward selection and the AIC. There is scope to include more detailed predictors later, see Whan and Schmeits 2018, but for this research we simply require a base model that improves upon the performance of the raw ensemble.^[I created a simulated example where overfitting was a concern and the correct predictors were not always selected. I need to explore this further]

## Climatology 

Belongs in a separate section but needs to be done

## Calibration

For an initisalisation time of 00:00 and the Winter season, a selection rank histograms are shown. The lead times are given by the facet labels on the y-axis. The calibration of the post-processed data is much improved compared with the raw ensemble.

```{r echo = F}
save_dir <- getwd()
rank_hist <- readRDS(paste0(save_dir, "/../data/plot_rainfall_rank_histograms.rds"))
rank_hist
```

## Scoring 

To compare how well the distribution of the post-processed ensemble and the raw ensemble we have estimated the continous ranked probability scores (CRPS). The distribution is skew as the median (lower two lines) is not a consistent estimator for the mean (upper two lines).^[want to fix the legend for this] Based on the median, the post-processed ensemble consistently scores better than the raw for the lowest quartiles. The differences in the mean score are less pronounced, but improvement is still clearly visible in the post-processed ensemble at longer lead times. However, a greater spread of scores is visible in the upper 95% confidence interval. There is a diurnal cyclic present in the scores which could be incorporated into the post-processing. 

```{r echo = F}
save_dir <- getwd()
crps_plot <- readRDS(paste0(save_dir, "/../data/plot_rainfall_crps.rds"))
crps_plot
```

## Brier Skill Scores 

Brier Skill Scores By Lead Time

## Diebold-Mariano



